\bf A hierarchical failure-time model for observational data exhibiting infant-mortality and wearout failure modes

### 2017-09-24 paper/paper.tex
Paper and References Edits per Meeker Meeting on 9.20

In this paper we present a method for joint estimation of multiple lifetime distributions based on the Generalized Limited Failure Population (GLFP) model. This 5-parameter model for lifetime data accommodates lifetime distributions with multiple failure modes:  early failure due to ``infant mortality'' and failure due to wearout. We fit the GLFP model using a hierarchical modeling approach.  Borrowing strength across populations, our method enables estimation with uncertainty of lifetime distributions even in cases where the number of model parameters is larger than the number of observed failures.  Moreover, using our Bayesian method, comparison of different product brands is straightforward because estimation of arbitrary functionals are easily computable using draws from the joint posterior distribution of the model parameters.
\end{abstract}

@@

The structure of the paper is as follows. In Section~\ref{sec:Background}, we introduce our proposed method and potential applications and examine previous related work in the literature. In Section~\ref{sec:Data}, we describe the Backblaze Hard Drive data set, which is used throughout as a motivating example. Section~\ref{sec:GLFP model} introduces the GLFP model and illustrates its use with a subset of the Backblaze data. In section~\ref{sec:Hierarchical GLFP model} we discuss using hierarchical modeling of the GLFP model parameters to extend the GLFP model to multiple populations. In section~\ref{sec:Data analysis} we illustrate applying the hierarchical GLFP model to a real data set. First, we do model selection among four models of increasing complexity, using approximate leave-one-out cross-validation. Next, for the selected model, we assess the model fit using a graphical comparison of replicated data sets generated from the posterior distribution to the observed data. Finally, we present an evaluative comparison of populations taking into account practical considerations related to product lifetime. Finally, in section~\ref{sec:Discussion}, we discuss the strengths and limitations of our method and consider extensions and alternatives.

\section{Background}
\label{sec:Background}
In engineering applications, a product can often fail from one out of a set of possible malfunctioning components.  For example, a computer system can break if the mother board, disc drive or power supply stop working.  Circuit boards (CB) can fail due to a manufacturing defect or later as a result of wearout.  The general name for such products is a series system where the lifetime of the product is the minimum failure time across different components or risks \citep{nelson}.  A common assumption in series systems is the time to failure for each risk is statistically independent.  Thus, the overall reliability of a unit is modeled using the product rule across all risks.  Parameter estimation is straightforward if the cause of failure is known for each observation.  However, in many situations the cause of failure is unknown to the analyst. This is referred to in the literature as ``masking''.\\

Previous papers have employed various data assumptions and methodologies to model masked lifetime failure data.  When modeling computer system failures Reiser et al. assumed each observed failure came from a known subset of failure modes, and estimation was performed using a Bayesian approach \citep{reiser}.  Chan and Meeker labeled the cause of circuit board failures as infant mortality, unknown, or wearout based on the time of observed failures.  This helped identify parameters when using maximum likelihood (ML) estimation.  Extending Chan and Meeker's analysis, Basu et al. performed a Bayesian analysis with informative priors to better identify early versus late failure modes without making any assumptions about the cause of failure \citep{basu}.  Berger and Sun introduced the Poly-Weibull distribution where cause of failure is the minimum of a several Weibull distributions \citep{berger}.  More recently, Ranjan et al. considered a competing risk model for infant mortality and wearout as a mixture of Weibull and exponential failure distributions \citep{ranjan}.  Treating the unknown failure modes as incomplete data, an expectation maximization (EM) algorithm providing ML estimates was used, in addition to Bayesian estimation.

@@

$$ P(T \le t) = 1 - \exp \left[ \left( -\frac{t}{\alpha} \right)^\beta \right]. $$

@@

\begin{equation} P(T \le t) = 1 - \exp \left\{ -\exp \left[ \frac{ \log (t) - \mu }{\sigma} \right] \right\}\\
= \Phi_{SEV}\left[\frac{\log (t) - \mu}{\sigma}\right],
\end{equation}
where $\Phi_{SEV}$ is the cdf of the standard smallest extreme value distribution (SEV). Thus, the Weibull distribution is a log-location scale family, with $\log(T)$ following an SEV distribution with location parameter $\mu$ and scale parameter $\sigma$. With this in mind, the corresponding density function is
\begin{equation}
p(t) = \frac{1}{\sigma t}\exp\left\{\frac{\log(t) - \mu}{\sigma} - \exp \left[ \frac{\log(t) - \mu}{\sigma} \right]\right\}\\
=\frac{1}{\sigma t} \phi_{SEV}\left[\frac{\log(t) - \mu}{\sigma}\right]
\end{equation}

An advantage conferred by writing the model in this way is that the log of the $p^{th}$ quantile, $t_p$, is a linear function of the parameters, which allows an analyst to apply prior information in a straightforward way.  Using the inverse cdf of the SEV distribution,
\begin{equation}
\log(t_p) = \mu + \sigma \Phi_{SEV}^{-1}(p).
\end{equation}
Because $\Phi(0)\approx 0.632$, we have $\log{\eta} = \mu  =t_{0.63}$. When there is more prior knowledge about a lower quantile, as is often the case with high-reliability products, using this parameterization allows the analyst to put prior distributions directly on those quantities. In addition, in the presence of censoring, $p$ can be chosen to make $\log(t_p)$ and $\sigma$ approximately independent. Since the choice of independent priors is common in practice, this observation lends further credence to the use of these reparameterizations.

\subsection{GLFP model}
\label{subsec:GLFP model}
Let $F_1,F_2$ be Weibull distributions with parameters $(\mu_1,\sigma_1)$ and $(\mu_2, \sigma_2)$, respectively.
The Generalized Limited Failure Population model (GLFP) of \citet{chan} is defined as follows: Let $T \sim \op{GLFP}(\pi, \mu_1,\sigma_1,\mu_2,\sigma_2)$. Then
$$P(T \le t) = F(t) = 1 - (1-\pi\, F_{1}(t))(1 - F_{2}(t)),\, t>0,\, 0 < \pi < 1.$$

@@

\subsection{Censoring and Truncation}
A common feature of lifetime data is right-censoring. Due to time or budget constraints it is rare all units are observed until failure. If a unit has not yet failed when the data is analyzed it is considered right-censored.  In other words, right-censoring puts a lower bound on the failure time.  Lifetime data for products with long expected lifetimes, such as hard drives, have a natural tendency to exhibit right-censoring.

@@

As of the first quarter of 2016, Backblaze was collecting and reporting data on 63 different drive-models.  Some drive-models have been running since 2013 or before, while others were added at a later date.  Data have been reported on 75,297 different hard drives that are or where in operation.  The number of drives varies by drive-model; some drive-models have only have a service record for a single drive whereas the maximum number of daily service records for a single drive model is 35,860.  Figure 1 shows a scatterplot of the total observed running time in hundred of thousands hours versus the total number of failures for drive-models with at least 3 failures.  For model identification, a minimum of three failures was the criterion for hard drive brands to be included in our analysis.

@@

\subsection{Analysis of a Single Drive-Model}
\label{subsec:ex1}
To illustrate the GLFP model, we will present an analysis of the drive-model with the most observed failures, drive-model 14.  The Kaplan-Meier estimate plotted below suggests at least two failure modes.  The curve has a slow ramp up in failures till about 10,000 hours, increases from about 10,000 to 20,000 hours, and then starts to gradually flatten out.  In addition, computer hard drives are known to have a mixture of early and late failures \citep{chan}.  Therefore, there is both an empirical and theoretical justification to apply the GLFP model.

\begin{figure}[H]
\centering
  \includegraphics[width=.8\textwidth]{km14-prob}
  \caption{Kaplan Meier estimate for drive model 14.  Uncertainty band correspond to pointwise standard errors (using Greenwood's formula.)}
  \label{fig1}
\end{figure}

To estimate the GLFP we use a Bayesian approach, selecting proper prior distributions to improve identification of the model parameters, which also ensures a proper posterior distribution. We reparameterize the component Weibull distributions using the $50^{th}$ percentile for the infant mortality failure mode ($t_{0.5,1}$) and the $20^{th}$ percentile for the wearout failure mode ($t_{0.2,2}$). Although non-standard, for purposes of exposition, we will use diagonal braces ($\langle.,.\rangle$) to refer to $95$ percent central probability intervals, rather than the standard model parameters when specifying prior distributions.

@@

These priors put probability mass on a wide range of values for all model parameters, much larger than we would expect for a typical Weibull distribution.  Thus, we consider the priors to be relatively uninformative.

Table 1 gives the posterior median and 95\% credible intervals for the 5 GLFP parameters for drive model 14.  The parameter $p$ is an estimate of the proportion of drives susceptible to to infant mortality.  As expected, this fraction is small with a median value of 0.05.  The parameter estimates for the two Weibull distributions also match intuition.  The wearout failure mode puts posterior probability on a values of $\sigma_1$ greater than 1, which corresponds to a decreasing hazard.  Conversely, the credible interval for $\sigma_2$, the wearout mode, is strictly below 1, an increasing hazard.  The two means times to failure (on the log scale) are also well identified with the infant mortality failure mode having an earlier mean time to failure than the wearout mode.

\begin{table}[H]
\centering
\begin{tabular}{rrrr}
  \hline
 & 2.5\% & 50\% & 97.5\% \\ 
  \hline
$\pi$ & 0.03 & 0.05 & 0.10 \\ 
 $\sigma_1$ & 1.25 & 1.96 & 15.54 \\ 
  $\sigma_2$ & 0.48 & 0.65 & 0.78 \\ 
  $\mu_1$ & 7.46 & 8.06 & 8.76 \\ 
  $\mu_2$ & 10.12 & 10.13 & 10.14 \\ 
   \hline
\end{tabular}
\caption{Posterior 95\% Credible Intervals for the 5 GLFP parameters for Drive Model 14.}
\label{table:1}
\end{table}

Probability plotting is a simple method to assess the adequacy of (log)location-scale families of probability models to the data.  After properly transforming the axes of a plot, graphing an empirical estimate of fraction failing as a function of time along with simultaneous confidence bands provides a visual check for distributional goodness of fit.  We applied this method using the Kaplan-Meier non-parametric estimate of the empirical cdf \citep{kaplan}.  With left truncation, however, the standard Kaplan-Meier estimator is biased, so we used an adjusted version (see Appendix for a detailed description).

In Figure \ref{ex1-overlay}, we overlay the posterior median of the fitted GLFP model onto the adjusted Kaplan-Meier estimate with axes on the Weibull probability scale.  Pointwise uncertainty bands associated with each estimator are included. While the Weibull model is inadequate for these data, the GLFP model fits quite well, as it is able to capture the acceleration of the hazard rate between 8000 and 20,000 hours. The discrepancy in the right tail of the distribution suggests that the GLFP may overestimate the proportion failing in the right tail of the distribution. While this discrepancy may appear troubling, we suggest that the right tail is of less importance for electronic goods due to obsolescence.

\begin{figure}[H]
\centering
  \includegraphics[width=.8\textwidth]{km14-prob-plus}
  \caption{Estimated GLFP for drive-model 14 plotted on Weibull paper.  Dashed line corresponds to the median of the posterior draws; pointwise 90\% credible band is in red.  Solid line is a non-parametric estimate with pointwise 90\% uncertainty band in blue.}
  \label{ex1-overlay}
\end{figure}

@@

In order to describe the entire population, consisting of different but similar sub-populations, and because of the limited amount of data from many of the sub-populations, we model the sub-population specific parameters hierarchically, borrowing strength across the sub-populations.  Let
\begin{equation}
y_{ig} \ind \op{GLFP}\left( \pi_g, \mu_{1g}, \sigma_{1g}, \mu_{2g}, \sigma_{2g} \right),
\end{equation}
where $g=1,\ldots,G$ indexes the sub-populations.  The likelihood for the hierarchical GLFP is a function of $\bm{\theta_g} = (\mu_{1g}, \sigma_{1g}, \mu_{2g}, \sigma_{2g}, \pi_{g})$.  Assuming the lifetimes of all units are independent within and across groups, the likelihood for the data, $i=1,\dots,n_g$,  is given by
\begin{equation*}
L(\bm{\Theta})= \prod_{g=1}^{G} \prod_{i=1}^{n_{g}} \left[\frac{f(t_{ig};\bm{\theta_g})}{1-F(t_{ig}^L;\bm{\theta_g})}\right]^{1-c_{ig}} \left[ \frac{1-F(t_{ig};\bm{\theta_g})}{1-F(t_{ig}^L;\bm{\theta_g})} \right]^{c_{ig}}
\end{equation*}

@@

As discussed in section \ref{sec:Weibull parameterization}, we reparameterize the Weibull in terms of log quantile and scale since the data we encounter are more informative about the left side of their respective lifetime distribution. We truncate the distribution of $\sigma_{2g}$ at $1$, restricting the wearout failure mode to have an increasing hazard function. We choose to model these parameters (or suitable transformations) with normal distributions because many populations require strong regularization which the normal, with its exponentially decreasing tails, provides.

In principle, we prefer this full model because we tend to believe that every population has a distribution with a distinct set of parameters. In practice, however, we may consider restrictions to reduce the number of parameters if the data do not support their inclusion. For example, we may consider a model with a common $\sigma_{1}$, letting $\sigma_{1g}=\sigma_1$ for all $g=1,\ldots,G$. A practical strategy is to begin with a simple model and gradually add complexity, reevaluating model adequacy at each iteration. In section \ref{sec:Model Comparisons}, we illustrate this process, fitting a series of models with increasing complexity and evaluating their performance using a measure of leave-one-out predictive error.

\section{Analysis of the Backblaze hard drive data}
\label{sec:Data analysis}
\subsection{Modeling}
We considered four models, all based on \ref{eq:hier-model}, which are identified by which model parameters are held constant across drive-models. These are (from most to least restrictive):

@@

The set of model specifications were chosen based data, interpretation of the model, as well as estimation considerations.  Product specific parameters for the the wearout failure mode $(\mu_{2g},\sigma_{2g})$, and the proportion defective $(\pi_g)$, are considered as a means to account for heterogeneity across brands in the right tails of the failure distribution.  Going from a common model for all product brands and gradually increasing the complexity of the model, we end up at Model 4.  Model 4 allows for the probability of infant mortality as well as the shape and scale parameters for the 2nd failure mode to vary by product.

For all of the models we consider, the parameters for the infant mortality failure mode are held in common across products.  We found that there was often insufficient information to model these parameters hierarchically.  Moreover, assuming a common distribution for infant mortality provides a meaningful interpretation and comparison of $\pi_g$ and $\pi_{g'}$ ($g \neq g'$).

\subsubsection{Prior Distributions}
\label{sec:Prior Distributions}
To complete the full probability model, we need to select prior distributions for the parameters governing the hierarchical model. We select proper priors to ensure a proper posterior.

For Models 1-4, different sets of restrictions required different prior specifications, which were assigned as follows:

\paragraph{Model 1.} Constrain all drive-models to the same GLFP distribution. For this ``reduced" model we assume the same priors as in Example 1.

\paragraph{Model 2.} $t_{p,g}$ varies by drive-model. To help with model identifiably, we tighten the priors on the infant mortality failure mode:
$$ \pi \sim \op{Logit-normal}\langle 7.0\times 10^{-3},\, 2.6 \times 10^{-1} \rangle,$$
$$\sigma_1 \sim \op{Log-normal}\langle 1.4 \times 10^{-1},\, 7.1\rangle,$$ 
$$t_{p1} \sim \op{Log-normal} \langle 2.2 \times 10^1,\, 5.5 \times 10^{4} \rangle.$$

\paragraph{Model 3.} $t_{p,g}$ and $\sigma_{g2}$ vary by drive-model. Priors for constrained parameters are the same as for model 2.

\paragraph{Model 4.} $\pi_g$, $t_{p,g}$ and $\sigma_{g2}$ vary by drive-model. Priors for constrained parameters are the same as for model 2.

Where applicable prior distributions on hyperparameters are as listed below. We follow the recommendation of \citet{gelman2014bayesian} to use half-Cauchy priors on hierarchical scale parameters. As for the location hyperparameters, we choose weakly informative prior distributions centered around the corresponding prior mean for the non-hierarchical model.
% The prior for $\eta_\pi$ puts 95\% of prior mass on the interval $(0.006, 0.27)$ for the median proportion defective (What justification? Should maybe be lower.) For $\eta_{\sigma, 2}$, $95\%$ of prior mass is on values greater than $0.037$. Since $\sigma_d$ is the reciprocal of the Weibull shape parameter, this corresponds roughly to an assumption that the median Weibull shape parameter is less than $1/0.037 = 27$. The prior for $\eta_{t_{p_2},2}$ implies that the median 20th percentile for non-defective units is less than greater than 3 days and less than 24 years.

@@

\subsection{Computation}
\label{sec:Computation}
Each model was fit using the {\tt
  rstan} \citep{rstan} package in {\tt R} \citep{r}, which implements a variant of Hamiltonian Monte Carlo (HMC)
\citep{betancourt}. HMC jointly updates all model parameters by simulating energy preserving paths with random initial momentums along the posterior density. This is done to reduce autocorrelation and efficiently explore the posterior. Multiple chains were run for 1500 iterations after 1500 warmup iterations: 4 chains were run for Models 1,2 and 3 and 16 chains were run for Model 4. The Gelman-Rubin potential scale reduction factor was used to provide a check for adequate mixing of the multiple chains. Upon convergence, $\hat{R}$, converges to 1. The number of model parameters for the four model fits are . The  maximum values of $\hat{R}$ across all model parameters for Models 1 throught 4 (5, 50, 95 and 140 parameters, respectively) were 1.005, 1.028, 1.009 and 1.004. Other diagnostics provided by the software (tree-depth, and divergent transitions) were clean. Plots of the posterior draws were inspected for parameters with the fewest effective samples; these did not suggest features that were inadequately explored.

@@

\subsection{Model Selection}
\label{sec:Model Comparisons}
As we have discussed, while we prefer to allow all of the GLFP model parameters to be drive-model specific, there are practical issues with fitting the full model. For example, due to heavy left-truncation, there may be insufficient information for particular drive models to estimate all of the parameters.

>>>>

In this paper we present a method for joint estimation of multiple lifetime distributions based on the Generalized Limited Failure Population (GLFP) model. This 5-parameter model for lifetime data accommodates lifetime distributions with multiple failure modes:  early failures due to ``infant mortality'' and failures due to wearout. We fit the GLFP model using a hierarchical modeling approach.  Borrowing strength across populations, our method enables estimation with uncertainty of lifetime distributions even in cases where the number of model parameters is larger than the number of observed failures.  Moreover, using our Bayesian method, comparison of different product brands is straightforward because estimation of arbitrary functionals are easily computable using draws from the joint posterior distribution of the model parameters.
\end{abstract}

@@

The structure of the paper is as follows. Section~\ref{sec:Background} introduces the proposed method and potential applications and examines previous related work in the literature.  Section~\ref{sec:GLFP model} introduces the GLFP model and illustrates its use with a subset of the Backblaze data.  In Section~\ref{sec:Data}, we describe the Backblaze Hard Drive data set, which is used throughout as a motivating example. In section~\ref{sec:Hierarchical GLFP model} we discuss using hierarchical modeling of the GLFP model parameters to extend the GLFP model to multiple populations. Section~\ref{sec:Data analysis} applies the hierarchical GLFP model to the hard disk drive data. First, we do model selection among four models of increasing complexity, using approximate leave-one-out cross-validation. Next, for the selected model, we assess the model fit using a graphical comparison of replicated data sets generated from the posterior distribution to the observed data. Finally, we present an evaluative comparison of populations taking into account practical considerations related to product lifetime. Finally, in section~\ref{sec:Discussion}, we discuss the strengths and limitations of our method and consider extensions and alternatives.

\section{Background}
\label{sec:Background}
In engineering applications, a product can often fail from one out of a set of possible malfunctioning components.  For example, a computer system can break if the mother board, disc drive or power supply stop working.  Circuit boards (CB) can fail due to a manufacturing defect or later as a result of wearout of certain components.  The general name for such products is a series system where the lifetime of the product is the minimum failure time across different components or risks \citep[Chapter 5]{nelson}.  A common assumption in series systems is the time to failure for each risk is statistically independent.  Thus, the overall reliability of a unit is modeled using the product rule across all risks.  Parameter estimation is straightforward if the cause of failure is known for each observation.  However, in many situations the cause of failure is unknown to the analyst. This is referred to in the literature as ``masking''.\\

Previous papers have employed various data assumptions and methodologies to model masked lifetime failure data.  When modeling computer system failures \citet{reiser} assumed each observed failure came from a known subset of failure modes, and estimation was performed using a Bayesian approach.  \cite{chan} labeled the cause of circuit board failures as infant mortality, unknown, or wearout based on the time of observed failures.  This helped identify parameters when using maximum likelihood (ML) estimation.  Extending Chan and Meeker's analysis, \citet{basu} performed a Bayesian analysis with informative priors to better identify early versus late failure modes without making any assumptions about the cause of failure.  \cite{berger} introduced the Poly-Weibull distribution where cause of failure is the minimum of a several Weibull distributions.  More recently, \cite{ranjan} considered a competing risk model for infant mortality and wearout as a mixture of Weibull and exponential failure distributions.  Treating the unknown failure modes as incomplete data, an expectation maximization (EM) algorithm providing ML estimates was used, in addition to Bayesian estimation.

@@

$$ \Pr(T \le t) = F(t) = 1 - \exp \left[ \left( -\frac{t}{\alpha} \right)^\beta \right]. $$

@@

\begin{equation} \Pr(T \le t) = F(t) = 1 - \exp \left\{ -\exp \left[ \frac{ \log (t) - \mu }{\sigma} \right] \right\}\\
= \Phi_{SEV}\left[\frac{\log (t) - \mu}{\sigma}\right],
\end{equation}
where $\Phi_{SEV}$ is the cdf of the standard smallest extreme value distribution (SEV). Thus, the Weibull distribution is a log-location scale family, with $\log(T)$ following an SEV distribution with location parameter $\mu$ and scale parameter $\sigma$. With this in mind, the corresponding density function is
\begin{equation}
f(t) =\frac{1}{\sigma t} \phi_{SEV}\left[\frac{\log(t) - \mu}{\sigma}\right]
\end{equation}

An advantage conferred by writing the model in this way is that the log of the $p$ quantile, $t_p$, is a linear function of the parameters, which allows an analyst to apply prior information in a straightforward way.  Using the inverse cdf of the SEV distribution,
\begin{equation}
\log(t_p) = \mu + \sigma \Phi_{SEV}^{-1}(p).
\end{equation}
Because $\Phi(0)\approx 0.632$, we have $\log{\eta} = \mu  \cong t_{0.63}$. When there is more prior knowledge about a lower quantile, as is often the case with high-reliability products, using this parameterization allows the analyst to put prior distributions directly on those quantities. In addition, in the presence of censoring, $p$ can be chosen to make $\log(t_p)$ and $\sigma$ approximately independent. Since the choice of independent marginal priors is common in practice, this observation lends further credence to the use of these reparameterizations.

\subsection{GLFP model}
\label{subsec:GLFP model}
Let $F_1,F_2$ be Weibull distributions with parameters $(\mu_1,\sigma_1)$ and $(\mu_2, \sigma_2)$, respectively.
The Generalized Limited Failure Population model (GLFP) of \citet{chan} is defined as follows: Let $T \sim \op{GLFP}(\pi, \mu_1,\sigma_1,\mu_2,\sigma_2)$. Then
$$\Pr(T \le t) = F(t) = 1 - (1-\pi\, F_{1}(t))(1 - F_{2}(t)),\, t>0,\, 0 < \pi < 1.$$

@@

\subsection{Censoring and Truncation}
A common feature of lifetime data is right-censoring. In the analysis of reliability field data, it is rare all units are observed until failure. If a unit has not yet failed when the data are analyzed it is considered right-censored.  In other words, right-censoring puts a lower bound on the failure time.

@@

As of the first quarter of 2016, Backblaze was collecting and reporting data on 63 different Drive-models.  Some Drive-models have been running since 2013 or before, while others were added at a later date.  Data have been reported on 75,297 different hard drives that are or where in operation.  The number of drives varies by Drive-model; some Drive-models have only have a service record for a single drive whereas the maximum number of daily service records for a single drive model is 35,860.  Figure 1 shows a scatterplot of the total observed running time in hundred of thousands hours versus the total number of failures for Drive-models with at least 3 failures.  For model identification, a minimum of three failures was the criterion for hard drive brands to be included in our analysis.

@@

\subsection{Analysis of a Single Drive-Model}
\label{subsec:ex1}
To illustrate the GLFP model, we will present an analysis of the drive-model with the most observed failures, Drive-model 14.  The Kaplan-Meier estimate plotted below suggests at least two failure modes.  The curve has a slow ramp up in failures till about 10,000 hours (early failures), increases more rapidly from about 10,000 to 20,000 hours (presumably wearout failures).  In addition, computer hard drives are known to have a mixture of early and late failures \citep{chan}.  Therefore, there is both an empirical and theoretical justification to apply the GLFP model.

\begin{figure}[H]
\centering
  \includegraphics[width=.8\textwidth]{km14-prob}
  \caption{Kaplan Meier estimate for Drive-model 14.  Uncertainty band correspond to pointwise standard errors (using Greenwood's formula.)}
  \label{fig1}
\end{figure}

To estimate the GLFP we use a Bayesian approach, selecting proper prior distributions to improve identification of the model parameters, which also ensures a proper posterior distribution. We reparameterize the component Weibull distributions using the $0.50$ quantile for the infant mortality failure mode ($t_{0.5,1}$) and the $0.20$ quantile for the wearout failure mode ($t_{0.2,2}$). \\

When eliciting prior distribution information it is much easier to ask about recognizable characteristics of a distribution instead of the parameters of the distribution.  Following the approach used in \cite[Section 15.2.2]{intervals} we will use diagonal braces ($\langle.,.\rangle$) to refer to $95$ percent central probability intervals, rather than the standard model parameters when specifying prior distributions.

@@

These prior distributions put probability mass on a wide range of values for all model parameters, much larger than we would expect for a typical Weibull distribution.  Thus, we consider the priors to be relatively uninformative.

Table 1 gives the posterior median and 95\% credible intervals for the 5 GLFP parameters for Drive-model 14.  The parameter $p$ is an estimate of the proportion of drives susceptible to to infant mortality.  As expected, this fraction is small with a median value of 0.05.  The parameter $\pi$ estimates for the two Weibull distributions also match intuition.  The infant mortality failure mode puts posterior probability on a values of $\beta_1$ less than 1, which corresponds to a decreasing hazard.  Conversely, the credible interval for $\beta_2$, the wearout mode, is strictly above 1, implying an increasing hazard function.  The two means times to failure (on the log scale) are also well identified with the infant mortality failure mode having an earlier mean time to failure than the wearout mode.

\begin{table}[H]
\centering
\begin{tabular}{rrrr}
  \hline
 & 2.5\% & 50\% & 97.5\% \\ 
  \hline
$\pi$ & 0.03 & 0.05 & 0.10 \\ 
 $\beta_1$ & 0.47 & 1.13 & 1.72 \\ 
  $\beta_2$ & 4.47 & 4.70 & 4.95 \\ 
  $\mu_1$ & 7.46 & 8.06 & 8.76 \\ 
  $\mu_2$ & 10.12 & 10.13 & 10.14 \\ 
   \hline
\end{tabular}
\caption{Posterior 95\% Credible Intervals for the 5 GLFP parameters for Drive-model 14.}
\label{table:1}
\end{table}

Probability plotting is a simple method to assess the adequacy of (log)location-scale families of probability models to the data.  After properly transforming the axes of a plot, graphing an empirical estimate of fraction failing as a function of time along with simultaneous confidence bands provides a visual check for distributional goodness of fit.  We applied this method using the Kaplan-Meier non-parametric estimate of the empirical cdf.  With left truncation, however, the standard Kaplan-Meier estimator is biased, so we used an adjusted version due to \cite[Chapter 11]{meeker} (see Appendix for a detailed description).

In Figure \ref{ex1-overlay}, we overlay the posterior median of the fitted GLFP model onto the adjusted Kaplan-Meier estimate with axes on the Weibull probability scale.  Pointwise uncertainty bands associated with each estimator are included. While the Weibull model is inadequate for these data, the GLFP model fits quite well, as it is able to capture the acceleration of the hazard rate between 8000 and 20,000 hours. The discrepancy around 30,000 suggests that the GLFP may overestimate the proportion failing in the far right tail of the distribution.  While this discrepancy may appear troubling, we suggest that the right tail is of less importance for electronic goods due to obsolescence.

\begin{figure}[H]
\centering
  \includegraphics[width=.8\textwidth]{km14-prob-plus}
  \caption{Estimated GLFP for Drive-model 14 plotted on Weibull paper.  Dashed line corresponds to the median of the posterior draws; pointwise 90\% credible band is in red.  Solid line is a non-parametric estimate with pointwise 90\% uncertainty band in blue.}
  \label{ex1-overlay}
\end{figure}

@@

In order to describe the entire population, consisting of different but similar sub-populations, and because of the limited amount of data from many of the sub-populations, we model the sub-population specific parameters hierarchically, borrowing strength across the sub-populations.  Let
\begin{equation}
y_{ig} \ind \op{GLFP}\left( \pi_g, \mu_{1g}, \sigma_{1g}, \mu_{2g}, \sigma_{2g} \right),
\end{equation}
where $g=1,\ldots,G$ indexes the sub-populations.  The likelihood for the hierarchical GLFP is a function of $\bm{\theta_g} = (\mu_{1g}, \sigma_{1g}, \mu_{2g}, \sigma_{2g}, \pi_{g})$.  Assuming the lifetimes of all units are independent within and across groups, the likelihood for the data, $i=1,\dots,n_g$,  is given by
\begin{equation*}
L(\bm{\Theta})= \prod_{g=1}^{G} \prod_{i=1}^{n_{g}} \left[\frac{f(t_{ig};\bm{\theta_g})}{1-F(t_{ig}^L;\bm{\theta_g})}\right]^{1-c_{ig}} \left[ \frac{1-F(t_{ig};\bm{\theta_g})}{1-F(t_{ig}^L;\bm{\theta_g})} \right]^{c_{ig}}
\end{equation*}
where $t_{ig}$ is the end of the observed lifetime of unit $i$ in group $g$; $t_{ig}^L}$ is the left truncation time for unit $i$ in group $g$; and $c_{ig}$ is an indicator if unit $i$ in group $g$ is right censored.  All times are again in hours.

@@

As discussed in Section \ref{sec:Weibull parameterization}, we reparameterize the Weibull distribution in terms of log quantile and scale since the data we encounter are more informative about the left side of their respective lifetime distribution. We truncate the distribution of $\sigma_{2g}$ at $1$, restricting the wearout failure mode to have an increasing hazard function. We choose to model these parameters (or suitable transformations) with normal distributions because many populations require strong regularization which the normal, with its exponentially decreasing tails, provides.

In principle, we prefer this full model because we tend to believe that every population has a distribution with a distinct set of parameters. In practice, however, we may consider restrictions to reduce the number of parameters if the data do not support their inclusion. For example, we may consider a model with a common $\sigma_{1}$, letting $\sigma_{1g}=\sigma_1$ for all $g=1,\ldots,G$. A practical strategy is to begin with a simple model and gradually add complexity, reevaluating model adequacy at each iteration. In Section \ref{sec:Model Comparisons}, we illustrate this process, fitting a series of models with increasing complexity and evaluating their performance using a measure of leave-one-out predictive error.

\section{Analysis of the Backblaze hard drive data}
\label{sec:Data analysis}
\subsection{Modeling}
We considered four models, all based on (\ref{eq:hier-model}), which are identified by which model parameters are held constant across drive-models. These are (from most to least restrictive):

@@

The set of model specifications were chosen based data, interpretation of the model, as well as estimation considerations.  Product specific parameters for the the wearout failure mode $(\mu_{2g},\sigma_{2g})$, and the proportion defective $(\pi_g)$, are considered as a means to account for heterogeneity across brands in the right tails of the failure distribution.  Going from a common model for all product brands and gradually increasing the complexity of the model, we end up at Model 4.  Model 4 allows for the probability of infant mortality as well as the shape and scale parameters for the 2nd failure mode to vary by Drive-model.

For all of the models we consider, the parameters for the infant mortality failure mode are held in common across Drive-models.  We found that there was often insufficient information to model these parameters hierarchically.  Moreover, assuming a common distribution for infant mortality provides a meaningful interpretation and comparison of $\pi_g$ and $\pi_{g'}$ ($g \neq g'$).

\subsubsection{Prior Distributions}
\label{sec:Prior Distributions}
To complete the full probability model, we need to select prior distributions for the parameters governing the hierarchical model. We select proper prior distributions to ensure a proper posterior distributions.

For Models 1-4, different sets of restrictions required different prior distribution specifications, which were assigned as follows:

\paragraph{Model 1.} Constrain all Drive-models to the same GLFP distribution. For this ``reduced" model we assume the same prior distributions as in Example 1.

\paragraph{Model 2.} $t_{p,g}$ varies by Drive-model. To help with model identifiably, we tighten the priors on the infant mortality failure mode:
$$ \pi \sim \op{Logit-normal}\langle 7.0\times 10^{-3},\, 2.6 \times 10^{-1} \rangle,$$
$$\sigma_1 \sim \op{Log-normal}\langle 1.4 \times 10^{-1},\, 7.1\rangle,$$ 
$$t_{p1} \sim \op{Log-normal} \langle 2.2 \times 10^1,\, 5.5 \times 10^{4} \rangle.$$

\paragraph{Model 3.} $t_{p,g}$ and $\sigma_{g2}$ vary by drive-model. Priors for constrained parameters are the same as for Model 2.

\paragraph{Model 4.} $\pi_g$, $t_{p,g}$ and $\sigma_{g2}$ vary by drive-model. Priors for constrained parameters are the same as for Model 2.

Where applicable prior distributions on hyperparameters are as listed below. We follow the recommendation of \citet{gelman2014bayesian} to use half-Cauchy prior distributions on hierarchical scale parameters. As for the location hyperparameters, we choose weakly informative prior distributions centered around the corresponding prior mean for the non-hierarchical model.
% The prior for $\eta_\pi$ puts 95\% of prior mass on the interval $(0.006, 0.27)$ for the median proportion defective (What justification? Should maybe be lower.) For $\eta_{\sigma, 2}$, $95\%$ of prior mass is on values greater than $0.037$. Since $\sigma_d$ is the reciprocal of the Weibull shape parameter, this corresponds roughly to an assumption that the median Weibull shape parameter is less than $1/0.037 = 27$. The prior for $\eta_{t_{p_2},2}$ implies that the median 20th percentile for non-defective units is less than greater than 3 days and less than 24 years.

@@

\subsection{Computation}
\label{sec:Computation}
Each model was fit using the {\tt
  rstan} \citep{rstan} package in {\tt R} \citep{r}, which implements a variant of Hamiltonian Monte Carlo (HMC)
\citep{betancourt}. HMC jointly updates all model parameters by simulating energy preserving paths with random initial momentums along the posterior density. This is done to reduce autocorrelation and efficiently explore the posterior. Multiple chains were run for 1500 iterations after 1500 warmup iterations: 4 chains were run for Models 1, 2 and 3 and 16 chains were run for Model 4. The Gelman-Rubin potential scale reduction factor was used to provide a check for adequate mixing of the multiple chains. Upon convergence, $\hat{R}$, converges to 1.  The  maximum values of $\hat{R}$ across all model parameters for Models 1 throught 4 (5, 50, 95 and 140 parameters, respectively) were 1.005, 1.028, 1.009 and 1.004. Other diagnostics provided by the software (tree-depth, and divergent transitions) were clean. Plots of the posterior draws were inspected for parameters with the fewest effective samples; these did not suggest features that were inadequately explored.

@@

\subsection{Model Selection}
\label{sec:Model Comparisons}
As we have discussed, while we prefer to allow all of the GLFP model parameters to be Drive-model specific, there are practical issues with fitting the full model. For example, due to heavy left-truncation, there may be insufficient information for particular Drive-models to estimate all of the parameters.
